---
title: "GameData Analysis (Kevin Ayala)"
author: "Kevin Ayala"
date: "3/9/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(knitr)
```

Read in csv. Mini EDA
```{r}
getwd()
data <- read.csv("/Users/kevinlorenzoayala/downloads/testData (1).csv")
head(data)


#Checking out the data 

GameID_count <- data %>% distinct(GameID) %>% count()
GameID_count   #there are a total of 7 different games in the data set

First_session_ofgame <- data %>% 
  mutate(FirstSessionDate = as.Date(FirstSessionDate, '%m/%d/%y')) %>% 
  group_by(GameID) %>% 
  summarise(FirstSessionDate = min(FirstSessionDate)) 

First_session_ofgame #has the first day of game having a user. 

```
```{r}
First_session_ofgame <-First_session_ofgame %>% mutate(NboGamers = NA)

x <- data %>% 
  mutate(FirstSessionDate = as.Date(FirstSessionDate, '%m/%d/%y'))

for (i in 1:7){
First_session_ofgame$NboGamers[i] = x %>% 
  filter(FirstSessionDate == First_session_ofgame$FirstSessionDate[i]) %>% 
  filter(GameID == First_session_ofgame$GameID[i]) %>% 
  distinct(UserID) %>% 
  count()

z <- as.numeric(unlist(First_session_ofgame$NboGamers))
First_session_ofgame$NboGamers <- z
}
```

```{r}
First_session_ofgame

ggplot(aes(x=GameID),data=First_session_ofgame)+xlab("Game")+
theme(axis.text=element_text(angle=90))+geom_bar(aes(weight=NboGamers,fill=GameID))+
ggtitle("Unique Users on First Day of Game Sessions") 
#Game 6 had the strongest number of users on the First Day of user appearing

```
Making an assumption that the data was put in correctly, so that the date 1/8/20 is 1/08/20 and not 1/18/20. And that 20 is 2020

The table shows the earliest date of when a user first appeared in game. As well as the number of unique players.
The visualization is a representation of the previous table.


```{r}
First_session_ofgame <- First_session_ofgame %>% 
  mutate(ThirdSessionDate = FirstSessionDate + 3)

First_session_ofgame <-First_session_ofgame %>%
  mutate(NboGamers_3 = NA)

for (i in 1:7){
First_session_ofgame$NboGamers_3[i] = x %>% 
  filter(FirstSessionDate == First_session_ofgame$ThirdSessionDate[i]) %>%
  filter(GameID == First_session_ofgame$GameID[i]) %>% 
  distinct(UserID) %>% 
  count()

v <- as.numeric(unlist(First_session_ofgame$NboGamers_3))
First_session_ofgame$NboGamers_3 <- v
}

First_session_ofgame

ggplot(aes(x=GameID),data=First_session_ofgame)+xlab("Game")+
theme(axis.text=element_text(angle=90))+geom_bar(aes(weight=NboGamers_3,fill=GameID))+
ggtitle("Amount of Unique Users on Day 3 After First Player Appeared")
```

Question 1) Which Game has the best Day 1, Day 3 Retention respectively based on the data?

First will be using classic or N day retention. Decided this after reading this article here:
https://amplitude.com/blog/n-day-retention-for-mobile-games

I assume that Day 1 is the date a user first appeared in the game. Hence why Day 1 = First Session Date
And assume that Day 3 is 2 days after Day 1. And not 3 days after Day 1
```{r}
x <- x %>% 
  mutate(SessionDate = as.Date(SessionDate, '%m/%d/%y'))   
#converting date factor, to date format

x <- x %>% 
  mutate(Day_1 = FirstSessionDate, Day_3 = FirstSessionDate + 2) 
#adding Day 1, and Day 3 to each User from the time they first appeared in the game. 

day1_data <- x %>% rowwise() %>% 
  mutate(match_Day1 = ifelse(between(SessionDate, FirstSessionDate, FirstSessionDate), 1, 0)) 

day1_return <- day1_data %>% 
  group_by(GameID, UserID) %>% 
  count(match_Day1) 


#filter out users who has count of 2 or more, a count of 1 indicates
#the first time a user first appeared. A count of 2 or indicates that 
# a user returned to the app after first appearance.
day1.retention <- day1_return %>% filter(n >= 2, match_Day1 == "1") %>% 
  distinct(UserID) 

day1.retention <- day1.retention %>% group_by(GameID) %>% count()
colnames(day1.retention)[2] <- "active_users.day1"
day1.retention  
```
Total number of unique users who returned to the game at least once again on the same day (Day1) after their first log in/play session.  
```{r}
Total_users <- day1_return %>% 
  group_by(GameID) %>% distinct(UserID) %>% count()

colnames(Total_users)[2] <- "total_users"
Total_users 
#total users per game. 
```

```{r}
day1.retention <- left_join(day1.retention, Total_users)
day1.retention <- day1.retention %>% 
  mutate(day1_retentionrate = active_users.day1/total_users)

day1.retention  
```

```{r}
day1.retention %>% 
  filter(day1_retentionrate == max(day1.retention$day1_retentionrate))

```

The game with the highest day 1 retention rate is game 1 relative to its user population. However, when looking in the table above, game 4 has the most returned users but with lower retention rate. To get percentages we can multiply by 100. 

Thus 59.4% is the highest retention rate for Day 1. Will keep future retention rates in decimal form.

```{r}
day3_data <- x %>% rowwise() %>%
  mutate(match_Day3 = ifelse(between(SessionDate, Day_3, Day_3), 1, 0)) 
#adding counter, 1 if user played a session exactly 3 days after install.
#keeping between() function as its useful for in the future to 
#check retention between specified dates. 

day3_return <- day3_data %>% 
  group_by(GameID, UserID) %>%   #grouping by GameID, and User ID 
  count(match_Day3)              #counting times a user appeared in a game


day3.retention <- day3_return %>%
  filter(match_Day3 == '1', n >= 1) %>%
  distinct(UserID) 
#filter out users who has a count of 1 and appeared on Day3

day3.retention <- day3.retention %>% 
  group_by(GameID) %>% count()  #getting count users of users on Day 3 per game

colnames(day3.retention)[2] <- "active_users.day3"
day3.retention  


Total_users <- day1_return %>% 
  group_by(GameID) %>% 
  distinct(UserID) %>% count()

colnames(Total_users)[2] <- "total_users"
Total_users #total users per game

day3.retention <- left_join(day3.retention, Total_users)
day3.retention <- day3.retention %>% 
  mutate(day3_retentionrate = active_users.day3/total_users) #classic retention rate

day3.retention 

```
```{r}
day3.retention %>% filter(day3_retentionrate == max(day3.retention$day3_retentionrate))
```
The game with the highest retention rate on Day 3 is game 4. Retention rate here meaning that Game 4 had the highest rate of users returning to the game on exactly the third day after first appearing in the game. Game 4 also had the highest number of users returning on the 3rd day, this is due to game 4 having a bigger population of players.

Short answer to question 1. Game 1 had biggest retention rate on Day 1 relative to its population. Game 4 also had the biggest retention rate relative to its population on Day 3. 

2) For the Game which has the highest number of users, which cohort of users based on
date had the best Day 1 and Day 3 overall?

The game with the highest number of Users
```{r}
highest.user_game <- x %>% group_by(GameID) %>% 
  distinct(UserID) %>% 
  count() #distinct to get unique values

highest.user_game <- as.data.frame(highest.user_game)
most.activegame <- highest.user_game %>% filter(n == max(n)) 
#selecting max of unique users, game 4 has the most all time users
most.activegame
```
There are 524 unique Users in Game 4, the highest. 
 

Number of Users Per Game, Visualized
```{r}
ggplot(aes(x=GameID),data=highest.user_game)+xlab("Game")+
theme(axis.text=element_text(angle=90))+geom_bar(aes(weight=n,fill=GameID))+
ggtitle("Total Number of Users")
#confirms table from Part 1
```
Vizualization of unique users per the lifetime of the game, according to the data

```{r}
game_4 <- x %>% filter(GameID == 'game_4')
head(game_4)  #subset of Data, game 4
```


```{r}
#adding counter, when user was active/in session on Day 1 and if played game on Day 3. 
#1 if true
game_4 <- game_4 %>% rowwise() %>% 
  mutate(match_Day1 = ifelse(between(SessionDate, Day_1, Day_1), 1, 0)) 
game_4 <- game_4 %>% 
  rowwise() %>% mutate(match_Day3 = ifelse(between(SessionDate, Day_3, Day_3), 1, 0))
head(game_4)

ggplot(aes(x=SessionDate),data = game_4)+xlab("Time in Days")+
theme(axis.text=element_text(angle=90))+
  geom_bar(aes(),fill = "skyBlue") +
  ggtitle("Total Sessions played for Game 4") 
```

This plot shows the total game 4 sessions played per day. The graph represents a total of new users and old users per day (up until that date) and how many total game sessions happened on that day. Assuming more total sessions played indicates there are more players.

```{r}
ggplot(aes(x=SessionDate, group = GameID),data = x)+xlab("Time")+
theme(axis.text=element_text(angle=90))+
  geom_bar(aes(fill=GameID))+
  ggtitle("Total Game Sessions Played") 
```
This graph shows the total sessions played per game per day. January 9, 2020 seems to be the most active day to be gaming in either Game 2 or game 4. It is interesting to note that game 2 was played mostly during January 9, 2020. Important to note that counts per game are added onto each other.


```{r}
#Counting Unique Users who got on the game again on Day 1
day_1_return.users <- game_4 %>% 
  group_by(FirstSessionDate, UserID) %>% 
  count(match_Day1) %>% 
  filter(match_Day1 == '1', n >= 2)
#we filter with n >= 2 because its Day 1. 


#Grouping and counting unique UserID per FirstSessionDate in final output
day_1_return.users <- day_1_return.users %>% 
  group_by(FirstSessionDate) %>% 
  count()


day_1_return.users <- as.data.frame(day_1_return.users)
colnames(day_1_return.users)[1] <- "Cohort"
colnames(day_1_return.users)[2] <- "frequency"  #renaming n

day_1_return.users %>% filter(frequency == max(frequency)) 
#getting max number of new users/player per date. 

head(day_1_return.users)
```
The highest number of new users who returned to game 4 on day 1 is 57 on January 09, 2020

```{r}
#Counting Unique Users who got on the game again on Day 3
day_3_return.users <- game_4 %>% 
  group_by(FirstSessionDate, UserID) %>% 
  count(match_Day3) %>% 
  filter(match_Day3 == '1', n >= 1) 

day_3_return.users <- day_3_return.users %>% 
  group_by(FirstSessionDate) %>% 
  count()

day_3_return.users <- as.data.frame(day_3_return.users)
colnames(day_3_return.users)[1] <- "Cohort"
colnames(day_3_return.users)[2] <- "frequency"
day_3_return.users %>% filter(frequency == max(frequency))
day_3_return.users
```
The highest number of players/users that returned to the game at exactly 3 days after they first appeared in the game is 10. This is a small number due to dataset size. 

As a gamer, it is my assumption that if someone still plays a game after/on 3 days, it is because they enjoy it and are likely to stick with it long term. Since we used exactly three days after and not within 3 days, there is the disadvantage of missing players who hypothetically could not play that specific day in the real world, even though they may turn to be a long time fan of the game and played at day 2 or any day after day 3.  

The best cohort for Day 3 on Classical Retention is on January 10, 2020, for Game 4

The best cohort for Day 1 and Day 3 is the cohort who first appeared in the game on January 09, 2020 and January 10, 2020 for Game 4

Vizualizing the above results

```{r}
cohort_day1 <- game_4 %>% group_by(FirstSessionDate, UserID) %>%
  count(match_Day1) %>% 
  filter(match_Day1 == '1', n >= 2) #data for visualization



ggplot(aes(x=FirstSessionDate), data = cohort_day1)+xlab("Time in Days")+
theme(axis.text=element_text(angle=90))+
  geom_freqpoly(aes(), color = "red")+
  ggtitle("Cohorts Who Returned to Game 4 On Day 1")  
```


```{r}
cohort_day3 <- game_4 %>% 
  group_by(FirstSessionDate, UserID) %>% 
  count(match_Day3) %>% 
  filter(match_Day3 == '1', n >= 1) #data for vizualization

ggplot(aes(x=FirstSessionDate), data = cohort_day3)+xlab("Time in Days")+
theme(axis.text=element_text(angle=90))+
  geom_freqpoly(aes(), color = "blue")+
  ggtitle("Cohorts Who Returned to Game 4 On Day 3")  
```


We see that the cohort who did best on Day 1, is the cohort who first appeared in the game on Jan. 09, 2020. 

The cohort that did best on Day 3, is the cohort who first appeared in the game on Jan. 10, 2020. Since Jan. 09, 2020 was the most popular day for the game based on the data, I suspect that players who first appeared on Jan. 10, 2020 are players who heard of the games popularity and became active users due to the presented popularity. 

I assume that the popularity is defined by the amount of new users per date, and that only frequent users (more than once) contribute to the games success/popularity. 

It is frequent users that are the ones who are mostly exposed to advertisments/microtransactions in game, thus from a busineess perspective, the cohort from Jan 09, 2020 and Jan 10, 2020 are the ones most likely to spend real money in game. 

